On The Impact of Atoms of Confusion in JavaScript Code

### 2021-01-29 4_Result.tex
revisao da secao de resultados

\begin{table*}[htbp]
\caption{Total number of correct answers.}
\label{tab:difference-correctness}
\centering{
\begin{tabular}{lccc|c} \toprule
 % & \multicolumn{2}{c}{Total Number of Correct Answers} & \\ [0.1cm]
  Atom & Confuse Code & Clean Code & $\Delta$(\%) & Effect Size (Odds Ratio) \\ \midrule
  Comma Operator &  28 &  65 & +132 & 19.02 \\ 
%   {\color{red}Automatic Semicolon Insertion} &  32 &  68 & +112 & 39.32 \\ 
  Post Increment &  48 &  64 & +33 & 4.83 \\ 
  Omitted Curly Braces &  47 &  58 &  +23 & 2.35 \\
  Assignment as Value &  56 &  68 &  +21 & 8.38 \\ 
  Implicit Predicate &  58 &  68 &  +17 & 6.94 \\ 
  Logic as Control Flow &  41 &  48 &  +17 & 1.53 \\ 
  Ternary Operator &  60 &  66 &  +10 & 2.73 \\ 
  Pre Increment &  50 &  53 &  +6 & 1.24 \\ 
  Arithmetic as Logic &  64 &  63 &  -2 & 0.84 \\ \bottomrule 
\end{tabular}
}
\end{table*}

@@

We first use the \emph{Pearson's Chi-squared test}
to investigate if there is a statistically significant difference
in the frequency of correct and incorrect answers---due to the versions
of the code snippets (confuse and clean code). The result indicates that the confusing version of the code snippets impacts code understanding (p-value $< 0.05$). We also use the \emph{Binomial Generalized Logistic Regression} analysis to investigate if either participant
\emph{education} or participant \emph{experience} impact on correctness. Although the results suggest that experience impact correctness, the education level does not. As such, we replicate the \emph{Pearson's Chi-squared test} for each experience group and found that the statistic difference is less significant for those developers with more than ten years of experience.
Finally, we measure the effect size of the clean version of the code snippets into the answers' correctness, using the \emph{Odds Ratio} (OR)
method (see the Effect Size column of Table~\ref{tab:difference-correctness}). We found a negligible effect size for the atom candidates Arithmetic as Logic, Logic as Control Flow, and Pre Increment. Nonetheless, for the remaining candidates, the effect size is significant. For instance, the Comma Operator leads to an \emph{Odds Ratio} of
\num{19.02}, meaning that the the odds of a correct answer are \num{19.02} times
higher when interpreting the clean version of a code---in comparison with the corresponding, confuse version of the code snippet.

@@

\begin{table*}[htbp]
\caption{Time in seconds to submit a {\color{red}correct} answer}
\label{tab:difference-time-taken}
\centering{
\begin{tabular}{lccc|cc} \toprule
Atom & Confuse Code & Clean Code & $\Delta$(\%) & Mann-Whitney test & Cliff's Delta \\ \midrule 
Comma Operator & 87.67 & 20.84 & -76.23 &  \colorbox{lightgray}{$<$ 0.01} & -0.53 \\ 
Logic as Control Flow & 108.94 & 51.07 & -53.12 &  \colorbox{lightgray}{$<$ 0.01} & -0.39 \\ 
% {\color{red}Automatic Semicolon Insertion} & 46.08 & 22.04 & -52.17 & 0.09 & -0.16 \\ 
Omitted Curly Braces & 48.85 & 30.00 & -38.58 & 0.92 & -0.01 \\ 
Implicit Predicate & 36.24 & 24.01 & -33.75 & \colorbox{lightgray}{$<$ 0.01} & -0.29 \\ 
Post Increment & 28.70 & 25.67 & -10.56 & 0.12 & 0.15 \\ 
Assignment as Value & 52.47 & 48.95 & -6.71 & 0.82 & 0.02 \\ 
Ternary Operator & 41.80 & 39.34 & -5.90 & 0.24 & -0.12 \\ 
Arithmetic as Logic & 28.82 & 37.20 & +29.06 & \colorbox{lightgray}{0.01} & 0.23 \\ 
Pre Increment & 30.71 & 42.45 & +38.19 & \colorbox{lightgray}{$<$ 0.01} & 0.27 \\ 
         \bottomrule
    \end{tabular}}
\end{table*}

\subsubsection*{Hypothesis Testing}
We use the non-parametric \emph{Mann-Whitney test} to
investigates the null hypothesis that developers 
spend the same amount of time to correctly
predict the outcome of a code snippet, regardless
of evaluating the confuse or the clean version
of the code. Table~\ref{tab:difference-time-taken}
shows the results, which suggest that we
should refute the null hypotheses for
five atom candidates (Comma Operator, Logic
as Control Flow, Implicit Predicate, Arithmetic
as Logic, and Pre Increment). For the first three,
the analysis suggest that the participants
need more time to predict the outcome of
the code snippet in the confuse code version. Interesting,
for the atom candidates Arithmetic as Logic and
Pre Increment, the participants take less time
to predict the outcome of the code snippets in
the confuse version.
We also computed the effect size using the Cliff's Delta method.
We found a large effect for the Comma Operator atom
candidate; and a moderate effect for the atom candidates
Logic as Control Flow, Implicit Predicate, Arithmetic
as Logic, and Pre Increment.

\begin{mh}
  The results suggest that developers spend
  significantly more time to predict the
  correct answer in the presence of the
  atom candidates Comma Operator, Logic as
  Control Flow, and Implicit Predicate. Conversely,
  the version of the code snippets with the
  atom candidates Arithmetic as Logic and
  Pre Increment demands less time to
  predict the correct answers. 
\end{mh}

@@

In this section we present the results of the 
interviews with the practitioners who agreed to participate
in the study. A total of 15 practitioners
took part of the interviews.
We collected information regarding programming
experience, familiarity with JavaScript, and their opinion
about the {\color{red}nine} atom candidates---by
(a) presenting them pairs of code snippets whose behavior were identical
and then (b) asking them to compare the code snippets readability.

@@

Supporting the results of the survey, Tables~\ref{tab:interview-results1}
and~\ref{tab:interview-results2} show that for twelve out of the
seventeen scenarios surveyed, the respondents prefer the version of the
code without the atom of confusion, and in no case the \emph{neutral}
ratio was higher than the option for the version without atoms of confusion.
%% An example entry is contained in Appendix~\ref{}. 
%% For a full listing of the code snippets, visit the paper repository.

\subsubsection*{Participants views of the atoms} Interesting,
only for the \emph{ternary operator} scenario the participants
prefer the version of the code with the atom of confusion,
instead of the more cleaner version without the atom of confusion.
Even in this case, some participants who opted for
the \emph{left-hand-side} (with the atom of confusion) version
still believed that the \emph{right-hand-side} (without the atom of
confusion) version was more readable (see Figure~\ref{code:ternary}). 
The following quotes were extracted from the transcripts with
three interviewees:

@@

The atom candidate Ternary Operator also opens up
the possibility for a derivative construct that JavaScript
allows which can be rather confusing, and that is the nested
ternaries construct, in which the right-hand side of the a
ternary operator can be another ternary construct.
While nested ternaries remove the number of lines that
would be necessary to construct using nested if-then-else statements,
they can become quite taxing, cognitively speaking, to understand.
Therefore, nested ternaries are a choice of atoms to be analyzed in future work.

@@

\rb{pela discuss\~{a}o anterior, parece-me que alguns
  desenvolvedores reconhecem que esses candidatos
  a \'{a}tomos introduzem certa dificuldade de compreens\~{a}o;
  mas mesmo assim optam por usar a constru\c c\~{a}o correspondente.
  Acho que cabe um box aqui com essa discuss\~{a}o sobre isso.}

When analyzing the Logic as Control Flow atom candidate, one of
the interviewees gave an example of his own experience
on why one should avoid writing code using this construct:

@@

This is an interesting perspective, because such small, yet tricky, constructs should be avoided, since maintenance is one of the longer and most costly parts of the software development life cycle, and one cannot make any assumptions about the other programmers' experience.

% \subsection{Common Remarks From the Open-ended Questions}

@@

\rb{seria importante incluir uma cita\c c\~{a}o para a \'{u}ltima afirma\c c\~{a}o acima.}

@@

\rb{esse pr\'{o}ximo par\'{a}grafo eh interessante,
  se refor\c carmos que os desenvolvedores tem
  experi\^{e}ncia com a linguagem. caso contr\'{a}rio,
  parece que nao fizemos uma boa selecao de participantes.
  algo para discutirmos.}

Moreover, such abstractions might lead a developer to start to program
using the language without actually understanding some of its basic aspects.
For instance, four participants argued that the language is not typed, two  believed the language runs exclusively on browsers, and two other developers stated that the language is natively asynchronous, all of which are incorrect statements. For the multi-purpose language that is one of the most, if not the most used one by practitioners
~\cite{}, a lack of understanding of the languages core
features may be a problem worth investigating in future studies.

\rb{seria importante apresentar uma conclus\~{a}o em rela\c c\~{a}o a essa parte do estudo. Acredito que seja interessante discutir um pouco essa ``faca de dois gumes'' sobre a flexibilidade, mencionada abaixo. }

\adriano{Fiquei em duvida se vale citar os contras mais interessantes que os entrevistados falaram sobre JS: (1x) o jeito como ela lida com 'this', (que esta ligado a ela ser puramente orientada a objetos, sem classe); (2x)falta de precisao para tratamento de ponto flutuante, o que pode a tornar inadequada para tratamento de dados/ciencias puras; e a faca de dois gumes que e a flexibilidade da linguagem: muitos falavam isso como ponto positivo e negativo ao mesmo tempo. \todo[inline]{eu acho que podemos mencionar sim}}

\begin{table}[!htb]
    \centering
    \caption{Interviews Round 1 - Summary of participants' preferences for code snippets \textsc{with} and \textsc{without} atoms of confusion (\textsc{aoc}). Participants were also allowed to choose \textsc{neutral} when they thought both sides were equally readable.}
    \label{tab:interview-results1}
    \begin{tabular}{lrrr}\toprule
      & \multicolumn{3}{c}{\textsc{Preference (\%)}} \\
      \cmidrule(lr){2-4}
         \textsc{atom}           & \multicolumn{1}{c}{\textsc{with aoc}}
                                      &  \multicolumn{1}{c}{\textsc{without aoc}}
                                               & \multicolumn{1}{c}{\textsc{neutral}} \\ \midrule
         Arithmetic as Logic     & 0  & 93.33  & 6.67  \\
         Assignment as Value     & 20 & 60     & 20    \\
         Comma Operator          & 0  & 100    & 0     \\
         Ternary Operator        & 60 & 26.67  & 13.3  \\
         Implicit Predicate      & 20 & 73.33  & 6.67  \\
         Logic as Control Flow   & 20 & 60     & 20    \\
         Omitted Curly Braces    & 0  & 100    & 0     \\
         Post Increment          & 20 & 73.33  & 6.67  \\
         Pre Increment           & 40 & 46.67  & 13.33 \\ \midrule
         \textsc{overall}        & 20 & 70.37  & 9.63  \\
         \bottomrule
    \end{tabular}
\end{table}

@@

\rb{continuar desse ponto}

\subsection{Discussion}
\label{sec:discussion}

>>>>

\begin{table*}[htbp]
\caption{Total number of correct answers.}
\label{tab:difference-correctness}
\centering{
\begin{tabular}{lccc|c} \toprule
 % & \multicolumn{2}{c}{Total Number of Correct Answers} & \\ [0.1cm]
  Atom & Confuse Code & Clean Code & $\Delta$(\%) & Effect Size (Odds Ratio) \\ \midrule
  Comma Operator &  28 &  65 & +132 & 19.02 \\ 
%   {\color{red}Automatic Semicolon Insertion} &  32 &  68 & +112 & 39.32 \\ 
  Post Increment &  48 &  64 & +33 & 4.83 \\ 
  Omitted Curly Braces &  47 &  58 &  +23 & 2.35 \\
  Assignment as Value &  56 &  68 &  +21 & 8.38 \\ 
  Implicit Predicate &  58 &  68 &  +17 & 6.94 \\ 
  Logic as Control Flow &  41 &  48 &  +17 & 1.53 \\ 
  Ternary Operator &  60 &  66 &  +10 & 2.73 \\ 
  Pre-Increment &  50 &  53 &  +6 & 1.24 \\ 
  Arithmetic as Logic &  64 &  63 &  -2 & 0.84 \\ \bottomrule 
\end{tabular}
}
\end{table*}

@@

We first use the \emph{Pearson's Chi-squared test}
to investigate if there is a statistically significant difference
in the frequency of correct and incorrect answers---due to the versions
of the code snippets (confuse and clean code). The result indicates that the confusing version of the code snippets impacts code understanding (p-value $< 0.05$). We also use the \emph{Binomial Generalized Logistic Regression} analysis to investigate if either participant
\emph{education} or participant \emph{experience} impact on correctness. Although the results suggest that experience impact correctness, the education level does not. As such, we replicate the \emph{Pearson's Chi-squared test} for each experience group and found that the statistic difference is less significant for those developers with more than ten years of experience.
Finally, we measure the effect size of the clean version of the code snippets into the answers' correctness, using the \emph{Odds Ratio} (OR)
method (see the Effect Size column of Table~\ref{tab:difference-correctness}). We found a negligible effect size for the atom candidates Arithmetic as Logic, Logic as Control Flow, and
Pre-Increment. Nonetheless, for the remaining candidates, the effect size is significant. For instance, the Comma Operator leads to an \emph{Odds Ratio} of
\num{19.02}, meaning that the the odds of a correct answer are \num{19.02} times
higher when interpreting the clean version of a code---in comparison with the corresponding, confuse version of the code snippet.

@@

\begin{table*}[htbp]
\caption{Time in seconds to submit a correct answer}
\label{tab:difference-time-taken}
\centering{
\begin{tabular}{lccc|cc} \toprule
Atom & Confuse Code & Clean Code & $\Delta$(\%) & Mann-Whitney test & Cliff's Delta \\ \midrule 
Comma Operator & 87.67 & 20.84 & -76.23 &  \colorbox{lightgray}{$<$ 0.01} & -0.53 \\ 
Logic as Control Flow & 108.94 & 51.07 & -53.12 &  \colorbox{lightgray}{$<$ 0.01} & -0.39 \\ 
% {\color{red}Automatic Semicolon Insertion} & 46.08 & 22.04 & -52.17 & 0.09 & -0.16 \\ 
Omitted Curly Braces & 48.85 & 30.00 & -38.58 & 0.92 & -0.01 \\ 
Implicit Predicate & 36.24 & 24.01 & -33.75 & \colorbox{lightgray}{$<$ 0.01} & -0.29 \\ 
Post Increment & 28.70 & 25.67 & -10.56 & 0.12 & 0.15 \\ 
Assignment as Value & 52.47 & 48.95 & -6.71 & 0.82 & 0.02 \\ 
Ternary Operator & 41.80 & 39.34 & -5.90 & 0.24 & -0.12 \\ 
Arithmetic as Logic & 28.82 & 37.20 & +29.06 & \colorbox{lightgray}{0.01} & 0.23 \\ 
Pre-Increment & 30.71 & 42.45 & +38.19 & \colorbox{lightgray}{$<$ 0.01} & 0.27 \\ 
         \bottomrule
    \end{tabular}}
\end{table*}

\subsubsection*{Hypothesis Testing}
We use the non-parametric \emph{Mann-Whitney test} to
investigates the null hypothesis that developers 
spend the same amount of time to correctly
predict the outcome of a code snippet, regardless
of evaluating the confuse or the clean version
of the code. Table~\ref{tab:difference-time-taken}
shows the results, which suggest that we
should refute the null hypotheses for
five atom candidates (Comma Operator, Logic
as Control Flow, Implicit Predicate, Arithmetic
as Logic, and Pre-Increment). For the first three,
the analysis suggest that the participants
need more time to predict the outcome of
the code snippet in the confuse code version. Interesting,
for the atom candidates Arithmetic as Logic and
Pre-Increment, the participants take less time
to predict the outcome of the code snippets in
the confuse version.
We also computed the effect size using the Cliff's Delta method.
We found a large effect for the Comma Operator atom
candidate; and a moderate effect for the atom candidates
Logic as Control Flow, Implicit Predicate, Arithmetic
as Logic, and Pre-Increment.

\begin{mh}
  The results suggest that developers spend
  significantly more time to predict the
  correct answer in the presence of the
  atom candidates Comma Operator, Logic as
  Control Flow, and Implicit Predicate. Conversely,
  the version of the code snippets with the
  atom candidates Arithmetic as Logic and
  Pre-Increment demands less time to
  predict the correct answers. 
\end{mh}

@@

In this section we present the results of the 
interviews with the practitioners who agreed to participate
in the study. A total of 15 practitioners
took part of the interviews.
We collected information regarding programming
experience, familiarity with JavaScript, and their opinion
about the nine atom candidates---by
(a) presenting them pairs of code snippets whose behavior were identical
and then (b) asking them to compare the code snippets readability.

@@

\subsubsection*{The Participants' perceptions of the atom candidates}
Supporting the results of the survey, Table~\ref{tab:interview-results1}
 shows that for eight out of the
nine scenarios surveyed, the maijority of the respondents agree that the version of the
code without the atom of confusion is easier to understand, and in no case the \emph{neutral}
ratio was higher than the option for the version without atoms of confusion.
%% An example entry is contained in Appendix~\ref{}. 
%% For a full listing of the code snippets, visit the paper repository. 
Only for the Ternary Operator atom candidate the participants
prefer the version of the code with the atom of confusion,
instead of the more cleaner version without the atom of confusion.
Even in this case, some participants who opted for
the \emph{left-hand-side} (with the atom of confusion) version
still believed that the \emph{right-hand-side} (without the atom of
confusion) version was more readable (see Figure~\ref{code:ternary}). 
The following quotes were extracted from the transcripts with
three interviewees:

@@

The atom candidate Ternary Operator also opens up
the possibility for a derivative construct that JavaScript
allows which can be rather confusing, and that is the nested
ternaries construct, in which the right-hand side of the a
ternary operator can be another ternary construct.
While nested ternaries remove the number of lines that
would be necessary to construct using nested if-then-else statements,
they can become quite taxing to understand.
Nested ternaries are a choice of atoms to be analyzed in future work.

@@

%% \rb{pela discuss\~{a}o anterior, parece-me que alguns
%%   desenvolvedores reconhecem que esses candidatos
%%   a \'{a}tomos introduzem certa dificuldade de compreens\~{a}o;
%%   mas mesmo assim optam por usar a constru\c c\~{a}o correspondente.
%%   Acho que cabe um box aqui com essa discuss\~{a}o sobre isso.}

When analyzing the Logic as Control Flow atom candidate, one of
the interviewees gave an example of his own experience
that might motivate one avoid writing code using this construct:

@@

This is an interesting perspective, because such small, yet tricky program constructs and idioms, might hinder novice developers in the task of maintaining code; and one cannot make any assumption about the programmers' experience of who is going to maintain a code in a long run.

\begin{table}[!htb]
    \centering
    \caption{Interviews Round 1 - Summary of participants' preferences for code snippets \textsc{with} and \textsc{without} atoms of confusion (\textsc{aoc}). Participants were also allowed to choose \textsc{neutral} when they thought both sides were equally readable.}
    \label{tab:interview-results1}
    \begin{tabular}{lrrr}\toprule
      & \multicolumn{3}{c}{\textsc{Preference (\%)}} \\
      \cmidrule(lr){2-4}
         \textsc{atom}           & \multicolumn{1}{c}{\textsc{with aoc}}
                                      &  \multicolumn{1}{c}{\textsc{without aoc}}
                                               & \multicolumn{1}{c}{\textsc{neutral}} \\ \midrule
         Arithmetic as Logic     & 0  & 93.33  & 6.67  \\
         Assignment as Value     & 20 & 60     & 20    \\
         Comma Operator          & 0  & 100    & 0     \\
         Ternary Operator        & 60 & 26.67  & 13.3  \\
         Implicit Predicate      & 20 & 73.33  & 6.67  \\
         Logic as Control Flow   & 20 & 60     & 20    \\
         Omitted Curly Braces    & 0  & 100    & 0     \\
         Post Increment          & 20 & 73.33  & 6.67  \\
         Pre-Increment           & 40 & 46.67  & 13.33 \\ \midrule
         \textsc{overall}        & 20 & 70.37  & 9.63  \\
         \bottomrule
    \end{tabular}
\end{table}

\subsubsection*{Confusion in JavaScript Code}

@@

@@

Moreover, such abstractions might lead a developer to start to program
using the language without actually understanding some of its basic aspects.
For instance, four participants argued that the language is not typed, two  believed the language runs exclusively on browsers, and two other developers stated that the language is natively asynchronous, all of which are incorrect statements. For the multi-purpose language that is one of the most, if not the most used one by practitioners,
a lack of understanding of the languages core
features may be a problem worth investigating in future studies.

@@
